### 2018-08-09 atomic-swaps/atomic-swaps-paper/intro.tex
[atomic swaps] new figures and updates in the table

\section{Introduction}\label{sec:intro}
An atomic swap enables involving parties to exchange cryptocurrencies, tokens on or off-chain without the involvement of a third party. Atomicity guarantees that either both sides of the swap happen, or neither. The risk of one party losing their assets is minimized. The concept of atomic swaps, which provides safe and quick trades, is being discussed since 2013\cite{altcoin}. Atomic swaps can be on-chain or off-chain. Off-chain swaps are instant, private and they almost doesn't incur any fees. Both type of swaps utilize hashed-time-locked-contracts (HTLCs) to guarantee that a party can't take any funds before providing his fund.

\section{Related Work}
Decred realized \textbf{the first on-chain atomic swap} between Decred and Litecoin~\cite{decred}. The need to mine new blocks to show the change of ownership can make the Decred's system slower. \textbf{The first on-chain Ethereum-Bitcoin atomic swap} is realized by Altcoin.io(REF!!). Lightning Labs performed \textbf{the first off-chain atomic swap} over the lightning network. Lightning network enables the off-chain exchange of coins  by creating payment channels between the trading parties. After the transaction is completed, the chains are updated and the channels are closed. Lightning network enables fast and scalable transactions and the involving parties don't have to pay transaction fees for every transaction.

%https://medium.com/@bitcoinatom/why-are-atomic-swaps-better-than-trading-on-exchanges-798492e150fd
%Their next goal: atomically swap any token for any other token, regardless of the blockchain it resides on.
Ethereum’s Raiden Network is analogous to Bitcoin’s Lightning Network and it provides instant, low-fee and scalable exchanges~\cite{Raiden}. The atomic swap of ERC20 tokens are performed on Raiden network.%Problem in decentralised exchanges: sacrificing transaction speed for settlement on-chain.
Zhang et al. propose Republic Protocol, which is a decentralized open-source dark pool protocol that provides atomic swaps of cryptocurrencies across Bitcoin and Ethereum blockchains~\cite{zhang2017republic}.%Republic: a  facilitating atomic swaps between cryptocurrency pairs across the . Trades are on a hidden orderbook and the matching is done using secure multiparty computation. It is a secure, decentralized, scalable dark pool protocol capable of handling billions in trading volume daily.
Their proposed system enables the exchange of Ether, ERC20 and Bitcoin over a decentralized dark pool. %execution without exposing price and volume, no trusted intermediary to operate a dark pool (dark pool = private exchanges where financial assets and instruments are traded and matched by an engine running on a hidden order book)Republic Protocol removes the risk of asset theft, confiscation or possibility of interference from a malicious exchange operator. Components: Decentralized hidden order book, Decentralized order matching, Atomic swap infrastructure, REN token nodes have no incentive to ignore an order, especially since they do not know the identity of the trader, nor the details of the order
In an atomic cross-chain swap setting, involving parties perform exchange across different blockchains like exchanging Bitcoin for Ether. In ~\cite{herlihy2018atomic}, a cross-chain swap is modeled as a directed graph where vertices correspond to the parties in the protocol and the arcs represent the transfers between them.
%An atomic swap protocol can be thought of as a trust-free, Byzantine-hardened form of distributed commitment. An atomic cross-chain swap is a special case of a distributed atomic transaction, although not all atomic transactions can be expressed as cross-chain swaps. If swaps are recurrent, then it would be useful to conduct swaps off-chain as much as possible, similar to the way that Lightning and Raiden networks support off-chain transactions for bitcoin and ERC20 tokens.FAIR EXCHANGE? The fair exchange problem ~\cite{franklin1998secure}, ~\cite{micali2003simple} is a precursor to the atomic cross-chain swap problem. Alice has a digital asset Bob wants, and vice-versa, and at the end of the protocol, either Alice and Bob have exchanged assets, or they both keep their assets. In the absence of blockchains, trusted, or semi-trusted third parties are required, but roles of those trusted parties can be minimized in clever ways.

\section{Our Implementation}

\begin{itemize}
  \item ETH to ERC20 token atomic swap is implemented.
  \item ERC20 to ERC20 token atomic swap is implemented.
  \item One contract can perform several swaps. Each swap is assigned a swapID.
  \item Each swap has three states: OPEN, CLOSED or EXPIRED.
\end{itemize}

\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{approve}
\caption{ETH to ERC20 with approve}
\label{fig:withapprove}
\vspace{-10pt}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{withoutapprove}
\caption{ETH to ERC20-like without approve}
\label{fig:withoutapprove}
\vspace{-10pt}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=1\textwidth]{ERC20toERC223}
	\caption{ERC20 to ERC223}
	\label{fig:ERC20toERC223}
	\vspace{-10pt}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=1\textwidth]{ERC223toERC223}
	\caption{ERC223 to ERC223}
	\label{fig:ERC223toERC223}
	\vspace{-10pt}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=1\textwidth]{ETHtoERC223}
	\caption{ETH to ERC223}
	\label{fig:ETHtoERC223}
	\vspace{-10pt}
\end{figure}

\section{Comparison of Token Standards}
\subsection{ERC721}
non-fungible, ERC20 compatible
\subsection{ERC223}
ERC20 has the problem of "non-recovarable" coins
ERC20 needs to execute two functions to transfer coins: approve and transferFrom. ERC223 consumes half the amount of gas than the amount needed to transfer an ERC20 token.
ERC223 offers 3 improvements:
\begin{itemize}
  \item no lost tokens
  \item reject non-supported tokens
  \item energy saving
\end{itemize}

tokenFallback: the receiving contract has a chance to do some work.
\subsection{ERC777}

\section{Atomic swaps with different settings}
Atomic swaps can be realized using different kinds of tokens and there can be different variations depending on the function calls used or collateralization. Table~\ref{Table:Comparison of different swaps} illustrates the possible settings and compares the amount of transactions required for each case.

\begin{table}[h!]
\footnotesize
\centering
%\Huge
%\resizebox{\columnwidth}{!}{%
\begin{tabular}{ |c|c|c|c|c|c| }
  \hline
   maker & taker & class & order transactions (margin) & account (onetime) transactions (fixed) \\ \hline
   ERC20 & ETH & approve & 2 & 4 \\ \hline
   ERC20 & ERC20 & approve & 2 & 5 \\ \hline
   ERC20-like & ETH & w/o approve & 2 & 4 \\ \hline
   ERC20-like & ERC20-like & w/o approve & 2 & 5 \\ \hline
   ERC223 & ETH & hybrid & 2 & 4\\ \hline
   ERC20 & ETH & collateralized & &\\ \hline
   ERC20 & ETH & partially collateralized & &\\ \hline
   ERC20 & ERC223 & hybrid & 2 & 5 \\ \hline
   ERC223 & ERC223 & collateralized & 2 & 5\\ \hline
\end{tabular}
\caption {swaps}
%}
\end{table}\label{Table:Comparison of different swaps} 
\textbf{ERC20 to ETH:} This setting is described in Figure~\ref{fig:withapprove}. Somebody deploys the swap contract. For each atomic swap between two parties, somebody (who may or may not be one of the transacting parties) initiates a swap object within the swap contract. Token holder calls approve on the ERC20 contract with the swap contract as the target/allowance holder/spender. The token holder may approve the swap contract for multiple transactions by setting an allowance greater than necessary for one swap. The holder of the ETH calls a settle function on the swap contract, sending their ETH along with it. This triggers the swap and the swap contract calls transferFrom on the ERC20 contract, transferring the tokens from the token holder to the ETH holder. The swap contract at this stage also sends the ETH to the token holder. The atomic swap is now complete. Deploying the swap contract and approving on the ERC20 contract are considered as the fixed transactions as they do not need to be performed for every swap. The swap initialization and calling of the settle function, both within the swap contract, must be carried out for every swap. When taken with the fixed cost transactions, an atomic swap takes up to four transactions total.

\textbf{ERC20 to ERC20:} For the fixed transactions, we have 5 in total as the other token holder should also approve the amount of tokens to be spent.

\textbf{ERC20-like to ETH:} This setting is described in Figure~\ref{fig:withoutapprove}. For this setting, the approve function in the ERC20 interface is not used.

\textbf{ERC20-like to ERC20-like:}