### 2018-09-17 frontrunning/Latex/setup/table.tex
[frontrunning] minor fix to the table

% ------------------------------------------------------------------------------------------------------------------------------------------------%
% Top 25 Dapps Table:%

\begin{table}[t]
\centering
\begin{tabular}{|c|c|}
\hline
\textbf{DApp Category}    											& \textbf{DApp Names (Rank)} \\  \hline
\multirow{6}{*}{Exchanges} 										& IDEX (1) \\ 
															& \textit{ForkDelta, EtherDelta}(2) \\
															& \textit{Bancor}(7) \\
															& The Token Store(13) \\
															& LocalEthereum(14) \\
															& \textit{0x Protocol}(23) \\
															& Kyber (22) \\  \hline

\multirow{9}{*}{\shortstack{Crypto-Collectible \\ Games \\ (ERC 721)}}		& \textit{CryptoKitties}(3) \\  
															& Ethermon(4)\\
															& Cryptogirl (9)\\
															& Gods Unchained TCG (12)\\
															& Blockchain Cuties(15)\\
															& ETH.TOWN! (16)\\
															& 0xUniverse (18)\\
															& MLBCrypto Baseball (19)\\
															& HyperDragons(25)\\ \hline

\multirow{8}{*}{Gambling}											& \textit{Fomo3D} (5) \\  
															& DailyDivs (6) \\ 
															& FomoWar (10) \\
															& FairDapp (11)\\
															& Zethr(17) \\
															& dice2.win (20) \\ 
															& PoWH 3D (8)	 \\ 
															& Ether Shrimp Farm(21) \\  \hline

\multirow{1}{*}{Name Services}									& \textit{Ethereum Name Service}(24)  \\  \hline

\end{tabular}
\caption{\footnotesize{Top 25 DApps based on recent user activity from \texttt{Dappradar.com} on September 4th, 2018. Dapps that are represented in Italic format are analyzied and discussed in this paper.}\label{tab:top25dapps}}
\end{table}

%TODO: mention the ones in italic are the ones we talk about
% ------------------------------------------------------------------------------------------------------------------------------------------------%