### 2018-09-17 auditing/Latex/sections/00-abstract.tex
[auditing] template ready

% !TEX root = ../main.tex
\begin{abstract}

Abstract goes here.

\end{abstract}