### 2018-09-17 auditing/Latex/sections/01-intro.tex

% !TEX root = ../main.tex
\section{Introductory Remarks}

Blah blah blah.

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =

\paragraph{Scope.} 
Blah blah blah.

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =

%\paragraph{Contributions.} 
%Our primary contributions are as follows.
%
%\begin{enumerate}
%
%\item Blah blah blah.
%
%\item Blah blah blah. 
%
%\item Blah blah blah. 
%
%\end{enumerate}