### 2018-09-17 auditing/Latex/main.tex

\documentclass[envcountsect]{llncs}
\pagestyle{headings} 
\bibliographystyle{abbrv}

% = = = = = Packages = = = = = %

\input{setup/packages}
%\input{setup/macros}

% = = = = = Title = = = = = %

\begin{document}
\frontmatter
\mainmatter

\title{\Large \bf A first look at auditing in a Blockchain world}

%\author{
%	Erica Pimentel\inst{1},
%	Emilio Boulianne\inst{1},
%	Shayan Eskandari\inst{2},
%	Jeremy Clark\inst{2}
%	}
%
%\institute{
%	\textsuperscript{1}JSMB,
%	\textsuperscript{2}CIISE \\ Concordia University
%	}

\maketitle

% = = = = = Abstract = = = = = %

\input{sections/00-abstract}

% = = = = = Main Body = = = = = %

\input{sections/01-intro}
\input{sections/02-body}
% = = = = = Bibliography = = = = = %

\bibliography{bib/}
%\nocite{*}

% = = = = = End Notes = = = = = %

\clearpage
\appendix

\end{document}