### 2018-09-18 auditing/Latex/sections/01-intro.tex
[auditting] some overall paper structure in Latex

% !TEX root = ../main.tex
\section{Introductory Remarks}

Blah blah blah.

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =

%\paragraph{Scope.} 
%Blah blah blah.

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =

\paragraph{\textbf{Methodology.}}  Blah blah
%Our primary contributions are as follows.
%
%\begin{enumerate}
%
%\item Blah blah blah.
%
%\item Blah blah blah. 
%
%\item Blah blah blah. 
%
%\end{enumerate}

\paragraph{\textbf{Relevance.}} Blah blah
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =

\section{Preliminaries and Related Work}

%------------------------------------------------------------------------------------------------
\subsection{Primary Stakeholders in an Audit}

\paragraph{Firm Management.}  Blah blah

\paragraph{Auditor.}  Blah blah

\paragraph{Investors.}  Blah blah

\paragraph{Financial Regulators.}  Blah blah

%------------------------------------------------------------------------------------------------
\subsection{Cryptoassets and firms that hold them}

%------------------------------------------------------------------------------------------------
\subsection{Presentation of Cryptoassets}

\paragraph{Cryptoassets.}  Blah blah

\paragraph{ICOs.}  Blah blah

\paragraph{Legal Status.}  Blah blah

\