### 2018-09-20 auditing/Latex/sections/00-abstract.tex
[auditing] transition to Latex. till Section 3

% !TEX root = ../main.tex
\begin{abstract}

The emergence of cryptocurrencies, like Bitcoin, and blockchain technologies, like Ethereum, have matured to a point that commercial firms operating in this space have begun to seek financial audits. At the time of writing, auditing firms are hesitant to audit such firms for a variety of reasons with a common underlying theme: the blockchain market introduces novel, technically sophisticated, and risky propositions. In this paper, we critically analyze the purported roadblocks to auditing blockchain firms, using a cross-disciplinary approach to bring cryptographers and auditors on the same page.

\end{abstract}