### 2018-09-25 frontrunning/Latex/setup/table.tex

% !TEX root = ../main.tex

% ------------------------------------------------------------------------------------------------------------------------------------------------%
% Top 25 Dapps Table:%

\begin{table}[t]
\centering
\scriptsize
\begin{tabular}{|c|l|}

\hline
\textbf{DApp Category}    										& \textbf{Names (Ranking)} \\  \hline
\multirow{6}{*}{Exchanges} 										& IDEX (1) \\ 
															& \textbf{ForkDelta, EtherDelta} (2) \\
															& \textbf{Bancor} (7) \\
															& The Token Store (13) \\
															& LocalEthereum (14) \\
															& Kyber (22) \\ 
															& \textbf{0x Protocol} (23) \\ \hline

\multirow{9}{*}{\shortstack{Crypto-Collectible \\ Games \\ (ERC 721)}}		& \textbf{CryptoKitties} (3) \\  
															& Ethermon (4)\\
															& Cryptogirl (9)\\
															& Gods Unchained TCG (12)\\
															& Blockchain Cuties (15)\\
															& ETH.TOWN! (16)\\
															& 0xUniverse (18)\\
															& MLBCrypto Baseball (19)\\
															& HyperDragons (25)\\ \hline

\multirow{8}{*}{Gambling}											& \textbf{Fomo3D} (5) \\  
															& DailyDivs (6) \\																								& PoWH 3D (8)	 \\ 
															& FomoWar (10) \\
															& FairDapp (11)\\
															& Zethr (17) \\
															& dice2.win (20) \\ 
															& Ether Shrimp Farm (21) \\  \hline

\multirow{1}{*}{Name Services}									& \textbf{Ethereum Name Service} (24)  \\  \hline

\end{tabular}
\caption{\footnotesize{Top 25 DApps based on recent user activity from \texttt{DAppRadar.com} on September 4th, 2018. We discuss the DApps that are in bold.}\label{tab:top25DApps}}
\end{table}

%TODO: mention the ones in italic are the ones we talk about
% ------------------------------------------------------------------------------------------------------------------------------------------------%