### 2018-09-25 auditing/Latex/sections/00-abstract.tex
Audit clean up

% !TEX root = ../main.tex
\begin{abstract}

The emergence of cryptocurrencies (\eg Bitcoin) and block\-chain technologies (\eg Ethereum) have matured to a point that commercial firms operating in this space are seeking audits of their financial statements. At the time of writing, auditing firms are hesitant to audit such firms for a variety of reasons with a common underlying theme: the blockchain market introduces novel, technically sophisticated, and risky propositions. In this paper, we critically analyze the purported roadblocks to auditing blockchain firms, using a cross-disciplinary approach to bring cryptographers and auditors on the same page.

\end{abstract}