### 2018-09-26 frontrunning/Latex/sections/abstract.tex
[frontrunning] final round on abstract and intro

% !TEX root = ../main.tex

\begin{abstract}

We consider \textit{front-running} to be a course of action where an entity benefits from prior access to privileged market information about upcoming transactions and trades. Front-running has been an issue in traditional financial instrument markets since the 1970's. With the advent of the blockchain technology, front-running resurfaced in new forms we explore here, due to blockchain's decentralized and transparent nature.
In this paper, we consider instances of front-running across the top 25 decentral applications (DApps) deployed on Ethereum. Additionally, we carry out a detailed analysis of \textsf{Status.im} initial coin offering (ICO) and show evidence of abnormal miner's behaviour indicative of front-running token purchases. Finally, we introduce a framework for categorizing proposed solutions to front-running and examine the applicability to the identified issues.

\end{abstract}