### 2018-12-06 frontrunning/Latex/main.tex
changes

\documentclass[envcountsect]{llncs}
\pagestyle{headings} 
\bibliographystyle{abbrv}

% = = = = = Packages = = = = = %

\input{setup/packages}
\input{setup/macros}

\usepackage{draftwatermark}
\SetWatermarkAngle{90}
\SetWatermarkHorCenter{0.2\textwidth}
\SetWatermarkColor{blue}
\SetWatermarkText{Working Draft}

% = = = = = Title = = = = = %

\begin{document}
\frontmatter
\mainmatter

\title{SoK: Transparent Dishonesty. Front-running attacks on Blockchain. }
\author{Shayan Eskandari, Seyedehmahsa Moosavi, \and Jeremy Clark}

\institute{
	Gina Cody School of Engineering and Computer Science\\ Concordia University
	}

\maketitle

% = = = = = Abstract = = = = = %

\input{sections/abstract}

% = = = = = Main Body = = = = = %

\input{sections/paper}

% = = = = = Ack = = = = = %

%\subsubsection*{Acknowledgements.}
%The authors thank the Autorit\'e des March\'es Financiers (AMF) for sponsoring this research through the Education and Good Governance Fund (EGGF), as well as NSERC through a Discovery Grant.

% = = = = = Bibliography = = = = = %

\clearpage
\bibliography{bib/Frontrunning_Ref}
%\nocite{*}

% = = = = = End Notes = = = = = %

\input{sections/appendix}

\end{document}