### 2018-12-13 frontrunning/Latex/sections/abstract.tex
into squeeze again

% !TEX root = ../main.tex

\begin{abstract}

We consider \textit{front-running} to be a course of action where an entity benefits from prior access to privileged market information about upcoming transactions and trades. Front-running has been an issue in traditional financial instrument markets since the 1970's. With the advent of the blockchain technology, front-running has resurfaced in new forms we explore here, instigated by blockchain's decentralized and transparent nature.
In this paper, we draw from a scattered body of knowledge and instances of front-running across the top 25 most active decentral applications (DApps) deployed on Ethereum. Additionally, we carry out a detailed analysis of \textsf{Status.im} initial coin offering (ICO) and show evidence of abnormal miner's behaviour indicative of front-running token purchases. Finally, we map proposed solutions to front-running into useful categories. 
\end{abstract}

% Jeremy: Commenting out this, too opinionated

% Developers don't have the mindset to design DApps with front-running in mind. This is an attempt to bring forward the subject and make them more aware of these type of attacks.