### 2018-12-20 frontrunning/Latex/setup/table.tex

% !TEX root = ../main.tex

% ------------------------------------------------------------------------------------------------------------------------------------------------%
% Top 25 Dapps Table:%

\definecolor{UnitedNationBlue}{rgb}{0.30,0.53,1}
\definecolor{LightSteelBlue}{rgb}{0.69,0.77,0.87}
\definecolor{LightGrey}{rgb}{0.83,0.83,0.83}

\begin{table}[t]
\captionsetup{justification = centering, singlelinecheck = false}
{\caption{Top 25 DApps based on recent user activity from \texttt{DAppRadar.com} on September 4th, 2018. The DApps that are in bold are discussed in this paper.}\label{tab:top25DApps}}
\begin{tabular}{|c|c|c|l|}

\hline
\rowcolor{LightSteelBlue}
 \textbf{DApp Category}    										& \textbf{Names} & Rank \\  \hline
\multirow{6}{*}{Exchanges} 										& IDEX & 1 \\ 
															& \cellcolor{LightGrey} \textbf{ForkDelta, EtherDelta} & 2 \\
															& \cellcolor{LightGrey} \textbf{Bancor} & 7 \\
															& The Token Store & 13 \\
															& LocalEthereum & 14 \\
															& Kyber & 22 \\ 
															& \cellcolor{LightGrey} \textbf{0x Protocol} & 23 \\ \hline

\multirow{9}{*}{\shortstack{Crypto-Collectible \\ Games \\ (ERC-721~\cite{erc721})}}
															& \cellcolor{LightGrey} \textbf{CryptoKitties} & 3 \\  
															& Ethermon & 4\\
															& Cryptogirl & 9\\
															& Gods Unchained TCG & 12\\
															& Blockchain Cuties & 15\\
															& ETH.TOWN! & 16\\
															& 0xUniverse & 18\\
															& MLBCrypto Baseball & 19\\
															& HyperDragons & 25\\ \hline

\multirow{8}{*}{Gambling}										& \cellcolor{LightGrey} \textbf{Fomo3D} & 5 \\  
															& DailyDivs & 6 \\													& PoWH 3D & 8	 \\ 
															& FomoWar & 10 \\
															& FairDapp & 11\\
															& Zethr & 17 \\
															& dice2.win & 20 \\ 
															& Ether Shrimp Farm & 21 \\  \hline

\multirow{1}{*}{Name Services}									& \cellcolor{LightGrey} \textbf{Ethereum Name Service} & 24  \\  \hline

\end{tabular}
\vspace{1em}
\end{table}

%TODO: mention the ones in italic are the ones we talk about
% ------------------------------------------------------------------------------------------------------------------------------------------------%