### 2019-01-30 frontrunning/Latex/main.tex
[frontrunning] Fix affiliate + comments about cryptokitties

\documentclass[envcountsect]{llncs}
\pagestyle{headings} 
\bibliographystyle{abbrv}

% = = = = = Packages = = = = = %

\input{setup/packages}
\input{setup/macros}

\usepackage{draftwatermark}
\SetWatermarkAngle{90}
\SetWatermarkHorCenter{0.2\textwidth}
\SetWatermarkColor{blue}
\SetWatermarkText{Working Draft}

% = = = = = Title = = = = = %

\begin{document}
\frontmatter
\mainmatter

\title{SoK: Transparent Dishonesty: front-running attacks on Blockchain.}
\author{}

\author{
	Shayan Eskandari\inst{\ddag},
	Seyedehmahsa Moosavi\inst{\dag},
	Jeremy Clark\inst{\dag}
	}

\institute{
	\textsuperscript{\dag} Gina Cody School of Engineering and Computer Science\\ Concordia University \\
	\textsuperscript{\ddag} ConsenSys Diligence
	}

\maketitle

% = = = = = Abstract = = = = = %

\input{sections/abstract}

% = = = = = Main Body = = = = = %

\input{sections/paper}

% = = = = = Ack = = = = = %

%\subsubsection*{Acknowledgements.}
%The authors thank the Autorit\'e des March\'es Financiers (AMF) for sponsoring this research through the Education and Good Governance Fund (EGGF), as well as NSERC through a Discovery Grant.

% = = = = = Bibliography = = = = = %

\bibliography{bib/Frontrunning_Ref}
%\nocite{*}

% = = = = = End Notes = = = = = %

\input{sections/appendix}

\end{document}